Enabling the Use of GPU Virtualization in Cloud Environments

### 2015-10-08 paper.tex
mod
revisado por adrian

\abstract{The use of accelerators in order to reduce the execution time of some 
application has become more popular during last years. These devices increment 
the computational power of a node thanks to its parallel architecture.
This trend has led several cloud vendors such as Amazon Web Services or some 
middlewares such as OpenStack to add to their current facilities, or to allow, 
instances of virtual machines equipped with NVIDIA GPUs. 
Accomplishing this requirements needs that the guest hosts must be equipped with 
GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
The solution presented in this work is based on GPU virtualization and shareability 
in order to reach an equilibrium between supply and demand of accelerators from the users. 
Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
With this software configuration, these GPUs can be accessed from any VM getting rid 
of the necessity of having physical GPUs in the guest host. 
Moreover, our project studies the viability of these join by using a public cloud 
service configuration and develops a module for OpenStack middleware in order to 
add the needed support for the virtualized devices and the logic to manage them. 
The results show a feasible configuration which adds flexibility to the current and well-known 
cloud solutions. 
}

@@

\subsection{Amazon Web Services}
\label{sec:aws}
AWS~\cite{aws} is a public cloud computing provider
and it is composed by several cloud services such as 
 cloud-based computation, storage and other functionality 
that enables organizations and/or individuals to deploy
services and applications on an on-demand basis. 
These services, replace IT infrastructure for most companies, and provide agility and instant elasticity matching 
perfectly with their software requirements.

@@

\section{\uppercase{State of Art}}
\label{sec:state}
Our solutions is based on GPU virtualization and sharing resources in order
to reach a fair balance between supply and demand.
Several attempts of achieving what we are pursing have been
made, but none of them are so ambitious as our project. On
the one hand the work done in~\cite{younge2013enabling} allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node. With the implied limitations of restricting the nodes not
to use more GPUs than the hosted in the machine and not
to share idle GPUs to other machines, several solutions have been presented.
The solution presented by the project gVirtus~\cite{giunta2010gpgpu} does virtualize GPUs and make
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor used, so
does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. 
This one continues without being integrated in a Cloud Computing Manager, though its main
drawback is that the code of the applications must be modified
in order to be run in their virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. Our proposal uses a different 
approach employing virtual machines and a cloud computing environment to allow to share GPUs in a flexible way.
On the other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
it is only focussed on compute intensive HPC applications. 
The main idea of our proposal goes further and apart from
bringing solutions for all kind of HPC applications, it is aimed
to boost the cluster flexibility in the use of GPUs.

@@

\subsection{Managing remote GPUs with OpenStack}
The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig}) to a solution where a new shared service takes part into the architecture and is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.

@@

Moreover, the users are also responsible for deciding whether a GPU (or a pool) will be assigned to other instances. 
This behaviour is refereed as ``scope'' and determines that a group of instances is logically connected to a pool of GPUs.
Working with ``public scope'' (second row of Figure~\ref{fig2}) means that the GPUs of a pool can be used simultaneously by all the instances linked to it.
Again, the GPU pool can be made of ``exclusive'' GPUs or ``shared''.

@@

\subsection{User Interface}
Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
First of all, the Instance Launch Panel should be provided with a new field, where the user could assign a existent GPU pool, create a new one or leave the instance without accelerators of this kind.
As soon as the option ``New GPU Pool'' had chosen, would appear the proper fields for the pool configuration (see Figure~\ref{fig:ui-launch}).
Furthermore, a new panel with the existent GPUs would display all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).

@@

The first set, in charge of providing the cloud environment,
was composed of 3 nodes, each one equipped with a Intel Xeon E7420 quadcore processor at
2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
To deploy an infrastructure as a service cloud, we have used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
A fully-featured OpenStack deployment requires at least three nodes: one controller node that manages the compute nodes 
where the VMs are hosted, one network node that manages the logic
virtual network for the VMs, and one or more compute nodes that 
run the hypervisor and virtual machines.

@@

We have designed 6 different set-ups which can be divided
in 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
“Exclusive” mode will provide, at most, 4 accelerators. Whilst
the number of available GPU in “Shared” mode will depend
on the partition size. In this particular case, we have halved
the memory, so we can use up to 8 GPUs. For each group, we
have deployed virtual clusters of 1, 2 and 3 nodes where will
run the processes of the application. The instances are based
on the flavor “m1.medium” which determines VM of 2 cores
and 4096 MB of memory.

MCUDA-MEME~\cite{Liu2010} has been the application executed with
different configurations in order to test the set-ups. MCUDA-
MEME is an MPI application which is accelerated by GPUs
Besides, each of its processes must have a GPU to not abort 
So, the number of GPUs will determine the number of processes that we can launch.

>>>>

\abstract{The use of accelerators in order to reduce the execution time of some 
application has become more popular during last years. These devices increment 
the computational power of a node thanks to its parallel architecture.
This trend has led several cloud vendors such as Amazon Web Services or some 
middlewares such as OpenStack to add to their current facilities, or to allow, 
instances of virtual machines equipped with NVIDIA GPUs. 
Accomplishing this requirements needs that the guest hosts must be equipped with 
GPUs which could be barely utilized if a non GPU-enabled Virtual Machine is running in the host.
The solution presented in this work is based on GPU virtualization and shareability 
in order to reach an equilibrium between supply and demand of accelerators from the users. 
Hence, we propose to decouple from the nodes, real GPUs by using the virtualization technology called rCUDA. 
With this software configuration, the GPUs can be accessed from any VM getting rid 
of the necessity of having physical GPUs in the guest host. 
Moreover, our project studies the viability of these join by using a public cloud 
service configuration and develops a module for OpenStack middleware in order to 
add the needed support for the virtualized devices and the logic to manage them. 
The results show a feasible configuration which adds flexibility to the current and well-known 
cloud solutions. 
}

@@

\subsection{Amazon Web Services}
\label{sec:aws}
AWS~\cite{aws} is a public cloud computing provider
and it is composed by several cloud services such as 
 cloud-based computation, storage and other functionality 
that enables organizations and/or individuals to deploy
services and applications on an on-demand basis. 
These services replace IT infrastructure for most companies and provide agility and instant elasticity matching 
perfectly with their software requirements.

@@

\section{\uppercase{State of Art}}
\label{sec:state}
Our solutions is based on GPU virtualization and sharing resources in order
to reach a fair balance between supply and demand.
Several attempts of achieving what we are pursing have been
made, but none of them are so ambitious as our project. On
the one hand the work done in~\cite{younge2013enabling} allows the VM managed
by the hypervisor Xen to access the GPUs in the physical
node, with the implied limitations of restricting the nodes not
to use more GPUs than the hosted in the machine and not
to share idle GPUs to other machines.
The solution presented by the project gVirtus~\cite{giunta2010gpgpu} virtualizes GPUs and make
them accessible for any VM in the cluster. However, this kind
of virtualization strongly depends on the hypervisor used, so
does its performance. Another similar solution is presented in~\cite{diab2013dynamic} by the name of gCloud. 
This one continues without being integrated in a Cloud Computing Manager, though its main
drawback is that the code of the applications must be modified
in order to be run in their virtual-GPU environment. A runtime component to provide abstraction
and sharing of GPUs is presented in~\cite{becchi2012virtual}, which allows scheduling policies 
to isolate and share GPUs in a cluster for a set of applications. Our proposal uses a different 
approach employing virtual machines and a cloud computing environment to allow to share GPUs in a flexible way.
On the other hand, the work exposed in~\cite{jungpgpu} is more mature, however,
it is only focussed on compute intensive HPC applications. 
The main idea of our proposal goes further and apart from
bringing solutions for all kind of HPC applications, it is aimed
to boost the cluster flexibility in the use of GPUs.

@@

\subsection{Managing remote GPUs with OpenStack}
The main idea is to move from the original OpenStack architecture (see Subfigure~\ref{subfig:os-orig})
 to a solution where a new shared service takes part into the architecture and it 
is responsible for all related to GPU accelerators (see Subfigure~\ref{subfig:os-mod}).
This new service would bring more flexibility when managing GPUs and new working modes for GPGPU computation in the Cloud.
As it is shown in Figure~\ref{fig:internal}, we have altered the OpenStack original Dashboard with a new parser, 
which split the HTTP query in order to make use of both, the GPGPU API for everything related to GPUs, and the nova API for the rest of the computation. 
The new GPGPU Service has brought a new feature which brings more flexibility when managing GPUs, from now on, we can create ``GPU-pools''. 
A ``GPU-pool'' is a set of independent GPUs (logically disattached from the nodes) that can be assigned to one or more instances.

@@

Moreover, the users are also responsible for deciding whether a GPU (or a pool) will be assigned to other instances. 
This behaviour is refereed as ``scope'' and it determines that a group of instances is logically connected to a pool of GPUs.
Working with ``public scope'' (second row of Figure~\ref{fig2}) means that the GPUs of a pool can be used simultaneously by all the instances linked to it.
Again, the GPU pool can be composed by ``exclusive'' or ``shared'' GPUs.

@@

\subsection{User Interface}
Several modifications have been planned in the OpenStack Dashboard in order to deal with the new features.
First of all, the Instance Launch Panel should be provided with a new field, 
where the user could assign an existent GPU pool, create a new one or leave the instance without accelerators of this kind.
As soon as the option ``New GPU Pool'' is chosen, the proper fields for the pool configuration appears(see Figure~\ref{fig:ui-launch}).
Furthermore, a new panel with the existent GPUs displays all the information related to GPUs (see Figure~\ref{fig:ui-rgpus}).

@@

The first set, in charge of providing the cloud environment,
was composed of 3 nodes, each one equipped with an Intel Xeon E7420 quadcore processor at
2.13 GHz, and 16 Gbytes of DDR2 RAM memory at 667 MHz.
To deploy an infrastructure as a cloud service, we have used OpenStack Icehouse version and QEMU/KVM 0.12.1 as the hypervisors.
A fully-featured OpenStack deployment requires at least three nodes: one controller node that manages the compute nodes 
where the VMs are hosted, one network node that manages the logic
virtual network for the VMs, and one or more compute nodes that 
run the hypervisor and virtual machines.

@@

We have designed 6 different set-ups which can be divided
into 2 groups: “Exclusive GPUs” and “Shared GPUs”. The
“Exclusive” mode will provide, at most, 4 accelerators. Whilst
the number of available GPU in “Shared” mode will depend
on the partition size. In this particular case, we have halved
the memory, so we can use up to 8 GPUs. For each group, we
have deployed virtual clusters of 1, 2 and 3 nodes where will
run the processes of the application. The instances are based
on the flavor “m1.medium” which determines VM of 2 cores
and 4096 MB of memory.

MCUDA-MEME~\cite{Liu2010} has been the application executed with
different configurations in order to test the set-ups. MCUDA-
MEME is an MPI application which is accelerated by GPUs.
Besides, each of its processes must have a GPU to not abort 
so, the number of GPUs will determine the number of processes that we can launch.
